---
title: Plotting HiC matrices
author: "Alejandro Reyes"
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('GenomeMatrix')`"
abstract: >
  "In the analysis of HiC or similar types of interaction data, it is useful
  to plot matrices while rotating them by -45 degrees. My motivation to write 
  this function was to visualize HiC data along with other genomic tracks, 
  such as gene annotations and chromatin marks. "
output:
  BiocStyle::html_document:
    highlight: pygments
    toc: true
    fig_width: 5
vignette: >
  %\VignetteIndexEntry{Plotting rotated matrices}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r echo=FALSE, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE,
                      cache = TRUE,
                      dev = "png",
                      message = FALSE,
                      error = FALSE,
                      warning = TRUE)
```

# Quick start

The minimum necessary to plot a matrix is simply a matrix. The function `matrixPlotter()` will rotate and plot only the upper half of the matrix. We specify the parameter `heightProp=1/2` for the height of the plot (y-axis) to be one half of the x-axis. 

```{r quickstart}
library(GenomeMatrix)
set.seed(100)
mat <- matrix( rnorm( 100, 0, 1 ), ncol=10 )
matrixPlotter(mat, heightProp=1/2) 
```

This is an ugly plot, but in the following sections I show a more useful example, with HiC data.

# HiC data

Below, I use a publicly available HiC dataset to demonstrate the main function. By providing a matrix, a *GenomicRanges* object where each element corresponds to a row (or a column) of the matrix and specifying the right `zlim=` and `heightProp=` parameters, we can visualize TADs horizontally. 

```{r}
library(HiTC)
data(Dixon2012_IMR90, package="HiCDataHumanIMR90")
mat <- as.matrix( hic_imr90_40@.Data[[1]]@intdata )
granges <- hic_imr90_40@.Data[[1]]@xgi
matrixPlotter( log2(mat+1), granges, 
               plotGR=GRanges("chr1", IRanges( 50000000, 60000000 ) ), 
               zlim=5, heightProp=1/4) 
```

Note that the output is a *ggplot2* object. Thus it is straightforward to arrange this into a single panel with other *ggplot2* plots, such as gene annotations from the *ggbio* package, using packages like *ggpubr*.

